---
title: "SR parameters estimation for Whiting (using the steepness parameter) "
output: html_document
editor_options:
  chunk_output_type: console
---

<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = T, message = F, warning = F)
library(kableExtra)
```

```{r input_data, echo=F}
## Weight at age
W_age <- read.csv("SR data Whiting from report WGCSE/weight_age.csv", row.names = 1)
colnames(W_age) <- c("0", seq(1:7))
w_age_2016 <- t(W_age[18,])
colnames(w_age_2016) <- "weight"

## Number at age
N_age <- read.csv("SR data Whiting from report WGCSE/n_age.csv", row.names = 1)
colnames(N_age) <- c("0", seq(1:7))
N_age_2016 <- t(N_age[10,])
colnames(N_age_2016) <- "N"

## Mat at age
M_Mat_age <- read.csv("SR data Whiting from report WGCSE/M_Mat_age.csv")
M_Mat <- M_Mat_age[,-2]
```


## Input parameters
```{r input_data2, echo=F}
## weight, N, Mat at age
data <- data.frame(M_Mat, weight = w_age_2016, age = N_age_2016)
library(kableExtra)
kable(x = data, format = "pandoc", digits = 2, align = "c", row.names = F)
```

## Etimation of $S_0$

$$ S_0 = \sum_{a=0}^{T}N_aW_aMat_a $$ 

Where

- $S_0$ : the unfished spawing biomass
- $N_a$ : the number at age
- $W_a$ : the weight at age
- $Mat_a$ : the maturity at age

The unfished spawing biomass $S_0$ :  

```{r input_data3, echo=F}
## S0
S0 <- apply(data, 1, function(x) x[2]*x[3]*x[4])
S0 <- sum(S0)
kable(x = S0, col.names = "S0") 
```

****  
<br>

The Beverton-Holt spawner-recruit function expressed with steepness parameter (Mace and Doonan 1988)

$$ R = \frac{0.8R_0hS}{0.2S_0(1-h)+(h-0.2)S}$$ 

Where

- $R$ : the recrutement
- $R_0$ : the unfished recruitment
- $h$ : the steepness parameter defined as the proportion of unfished recruitment $R_0$ produced by 20% of unfished population fecundity or spawning biomass $S_0$
- $S$ : the spawing stock biomass

## Likelihood function

```{r LL}
load("species_data.RData")
WHG_data <- sp_data[[1]]

ssb <- WHG_data$SSB
rec <- WHG_data$recruitment
h <- 0.81

LL <- function(par) {
  R0 <- exp(par[1])
  c <- exp(par[2])
  
  R <- 0.8*R0*h*ssb / (0.2*S0*(1-h) + ssb*(h-0.2))
  
  ll = dlnorm(rec, meanlog = log(R)-(c^2)/2, sdlog = c, log = T)
  
  -sum(ll)
}

LL_opt <- nlminb(start = log(c(10000, 0.2)),  objective = LL)
LL_opt

R0_est <- exp(LL_opt$par[1])
```

## SR parameters
# The Beverton and Holt stock recruitment relationship

$$R=\frac{B}{\alpha + \beta B} $$
Where

- $R$ : the recruitement
- $B$ : the spawing stock biomass
- $\alpha$ : he inverse of the initial slope of the curve
- $\beta$ : the inverse of asymptotic recruitment 

$$\alpha = \frac{S_0}{R_0} \frac{1-h}{4h} $$ 
$$\beta = \frac{5h-1}{4hR_0} $$ 

Where

- $R_0$ : the recruitment when $F=0$
- $h$ : the steepness parameter defined as the proportion of unfished recruitment $R_0$ produced by 20% of unfished population fecundity or spawning biomass $S_0$
- $S_0$ : the spawing stock biomass when at $F=0$

(Equations from Mangel et al. 2009)

```{r SR, echo=F }
## S-R parameters
a <- (S0/R0_est) * (1-h)/(4*h)
b <- (5*h-1) / (4*h*R0_est)

sr1 <- function(x){ x / (a + b*x) }
```

```{r plot, echo=F}
s <- seq(0, 10e5, by = 1000)
plot(ssb, rec, bty = "l",  xlim = c(0, 1e5), ylim = c(0, 3e6),
     xlab = "ssb", ylab = "rec")
lines(s, sr1(s))
```


